### 2016-06-07 paper/icfp16-cameraready/related.tex
update related work

\paragraph{Diagnosing and Repairing Type Errors}
\label{sec:diagnosis-repair}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%
Thus, many groups have explored techniques to pinpoint the true source
of the error and recommend fixes.

Lerner \etal~\cite{Lerner2007-dt} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{Chen2014-gd} use a variational type
system to allow for the possibility of changing an expression's type,
and search for an expression whose type can be changed such that type
inference would succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types used in the program.
%
Neubauer and Thiemann~\cite{Neubauer2003-xv} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{Pavlinovic2014-mr} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis~\etal~\cite{Vytiniotis2012-gh} extend the \haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.
%
Bayne~\etal~\cite{Bayne2011-cn} also provide a semantics for running
ill-typed (\java) programs, but in constrast transform the program to
perform nearly all type checking at run-time. The key difference between
Bayne~\etal and our work is that we use the dynamic semantics to
automatically search for a witness to the type error, while their focus
is on incremental, programmer-driven testing.

@@

Instead of enumerating values, which may trigger the same path through
the program, one might enumerate paths. 
%
Dynamic-symbolic execution~\cite{Godefroid2005-am,Cadar2008-kg,Tillmann2008-qc}
%
combines symbolic execution (to track which path a given input
triggers) with concrete execution (to ensure failures are
not spurious). The system collects a path condition during execution,
which tracks symbolically what conditions must be met to trigger the
current path. Upon successfully completing a test run, it negates the
path condition and queries a solver for another set of inputs that
satisfy the negated path condition, \ie inputs that will not trigger the
same path. Thus, it can prune the search space
much faster than techniques based on enumerating values, but is limited
by the expressiveness of the underlying solver. Our operational
semantics is amenable to dynamic-symbolic execution, one would just need
to collect the path condition and replace our implementation of \gensym
by a call to the solver. We chose to use lazy, random generation instead
because it is efficient, and % does not incur the overhead of an external solver, and
produces high coverage for our domain of novice programs.

>>>>

\paragraph{Localizing and Repairing Type Errors}
\label{sec:diagnosis-repair}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%
Thus, many groups have explored techniques to pinpoint the true source
of the error and recommend fixes.

% \subparagraph{Localization}
The traditional Damas-Milner type inference
algorithm~\cite{Damas1982-uw} reports the first program location where a
type mismatch is discovered (subject to the traversal
strategy~\cite{Lee1998-ys}).
%
As a result the error can be reported far away from its
source~\cite{McAdam1998-ub} without enough information to guide the
user.
%
Type-error slicing~\cite{Haack2003-vc,Schilling2011-yf,Rahli2015-tt,Sagonas2013-bf,Gast2004-zd,Neubauer2003-xv}
%treats type inference as a constraint-satisfaction problem and
recognizes this flaw and instead produces a slice of the program
containing \emph{all} program locations that are connected to the type
error.
%
%% \citet{Neubauer2003-xv} present a decidable type system based on
%% discriminative sum types, in which all terms are typeable and type
%% derivations contain all type errors in a program. They then use the
%% typing derivation to slice out the parts of the expression related to
%% each error.
%
Though the program slice must contain the source of the error, it can
suffer from the opposite problem of providing \emph{too much}
information, motivating recent work in ranking the program locations.
%
Zhang~\etal~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit using Bayesian reasoning.
%
\citet{Pavlinovic2014-mr,Pavlinovic2015-kh} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.

% \subparagraph{Repairing}
In addition to localizing the error, \citet{Lerner2007-dt} attempt to
suggest a fix by replacing expressions (or removing them entirely) with
alternatives based on the surrounding program context.
%
\citet{Chen2014-gd} use a variational type system to allow for the
possibility of changing an expression's type, and search for an
expression whose type can be changed such that type inference would
succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types used in the program.
%
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}
%

@@

\paragraph{Improving Error Messages}
The content and quality of the error messages themselves has also been
studied extensively.
%
\citet{Marceau2011-ok,Marceau2011-cy} study the effectiveness of error
messages in novice environments and present suggestions for improving
their quality and consistency.
%
\citet{Hage2006-hc} identify a variety of general heuristics to improve
the quality of type error messages, based on their teaching experience.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} provide methods for library authors to specialize
type errors with domain-specific knowledge.
%
The difference with our work is more pronounced here as we do not
attempt to improve the quality of the error message, instead we search
for a witness to the error and explain it with the resulting execution
trace.

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
\citet{Vytiniotis2012-gh} extend the \haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.
%
\citet{Bayne2011-cn} also provide a semantics for running
ill-typed (\java) programs, but in constrast transform the program to
perform nearly all type checking at run-time. The key difference between
Bayne~\etal\ and our work is that we use the dynamic semantics to
automatically search for a witness to the type error, while their focus
is on incremental, programmer-driven testing.

@@

Instead of enumerating values, which may trigger the same path through
the program, one might enumerate paths. 
%
Dynamic-symbolic execution~\cite{Godefroid2005-am,Cadar2008-kg,Tillmann2008-qc}
%
combines symbolic execution (to track which path a given input triggers)
with concrete execution (to ensure failures are not spurious). The
system collects a path condition during execution, which tracks
symbolically what conditions must be met to trigger the current
path. Upon successfully completing a test run, it negates the path
condition and queries a solver for another set of inputs that satisfy
the negated path condition, \ie inputs that will not trigger the same
path. Thus, it can prune the search space much faster than techniques
based on enumerating values, but is limited by the expressiveness of the
underlying solver.

Our operational semantics is amenable to dynamic-symbolic execution, one
would just need to collect the path condition and replace our
implementation of \gensym by a call to the solver. We chose to use lazy,
random generation instead because it is efficient, and % does not incur
the overhead of an external solver, and produces high coverage for our
domain of novice programs.
